---
title: "Homework 3"
author: "Kairui Wang"
date: "2023-10-09"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

1. Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

2. Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r, message=FALSE}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

#### Read in the 'BRFSS' data.

```{r}
data("brfss_smart2010")
```

#### Do some data cleaning.

```{r}
brfss_df = 
  brfss_smart2010 |>
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  filter(response %in% c("Excellent","Very good","Good","Fair","Poor")) |>
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good","Very good","Excellent"))) |> 
  rename("state" = "locationabbr","location" = "locationdesc")

```

#### Answer of the questions

1. In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_df |> 
  filter(year == "2002") |>
  group_by(state) |>
  summarize(days = n_distinct(location)) |> 
  filter(days >= 7) |> 
  knitr::kable()
```

```{r}
brfss_df |> 
  filter(year == "2010") |>
  group_by(state) |>
  summarize(days = n_distinct(location)) |> 
  filter(days >= 7) |> 
  knitr::kable()
```

As shown in the table, 
In 2002, CT, FL, MA, NC, NJ, PA were observed at 7 or more locations.
In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA were observed at 7 or more locations.

2. Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_exc = 
  brfss_df |> 
  filter(response == "Excellent") |>
  group_by(state, year) |>
  mutate(ave_data = mean(data_value)) |>
  select(year, state, ave_data) 
ggplot(brfss_exc, aes(x = year, 
                  y = ave_data,
                  group = state, 
                  color = state)) +
  geom_line(aes(color = state, group = state), alpha = .5) +
  labs(
    title = "Average value over time for states", 
    x = "year", 
    y = "mean value") 
```

Comments
The mean value of each state are fluctuating along with the time, but all of them are between 10 and 30.

3. Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_df |>
  filter(year == 2006 | year == 2010) |>
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .5) +
  facet_grid(year ~ .) +
  labs(
    title = "Distribution of data_value", 
    x = "data_value")
```

Comments
The distribution of data value for a particular response category in 2006 and 2010 did not alter significantly. People who gave "Poor" responses typically scored around 4, while those who gave "Fair" responses scored mostly around 10, those who gave "Good" responses scored mostly around 20, those who gave "Very Good" responses scored mostly around 30, and those who gave "Excellent" responses scored mostly around 36.


## Problem 3

1. Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

#### tidy nhanes_covar

```{r}
nhanes_covar = 
  read_csv("nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
  drop_na() |>
  mutate( 
    sex = case_match(
      sex,
      1 ~ "male", 
      2 ~ "female"
    ), 
    education = case_match(education, 
                           1~ "Less than high school", 
                         2 ~ "High school equivalent", 
                         3 ~ "More than high school"))

nhanes_covar$education = factor(nhanes_covar$education, levels = c("Less than high school", "High school equivalent", "More than high school")) 
```

#### tidy nhanes_accel

```{r}
nhanes_accel = 
  read_csv("nhanes_accel.csv") |>
  janitor::clean_names()
```

#### combine the data set.

```{r}
nhanes_df = inner_join(nhanes_covar, nhanes_accel, by = "seqn")
```

There are `r nrow(nhanes_df)` observations and `r ncol(nhanes_df)` variables in the merged and filtered dataset.

2. Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
nhanes_df |> 
  group_by(education, sex) |>
  summarize(count = n()) |> 
  pivot_wider(names_from =  sex, 
              values_from = count) |> 
  knitr::kable()
```

```{r}
nhanes_df$education = factor(nhanes_df$education)

ggplot(nhanes_df, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distribution by Gender and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal()
```

Comments
At high school equivalent, the number of males is larger than female. For the less than high school education level, number of female is a little bit larger than male. 

Participants who have less than a high school diploma are often older than 70, whether they are male or female. 
Male participants with a high school diploma are typically around 60 years old, whereas female participants with a high school diploma are often approximately 74 years old. 
Male participants who have completed high school are typically around 31, whereas female participants who have completed high school are often around 33. 

The participant age decreases as education level increases.


3. Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}
nhanes_df$sum_nhanes = rowSums(nhanes_df[, 6:1445])

ggplot(nhanes_df, aes(x = age, y = sum_nhanes, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(.~ education) +
  labs(
    title = "Total activity time over age",
    x = "Age",
    y = "Total Activity"
  ) 
```

Comments
While the participant's activity time typically decreases with age, there may occasionally be one or two peaks at a given age. Male participants' fluctuation range is less than that of female participants. 

4. Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}
hour_activity = nhanes_df |>
  pivot_longer(cols = min1:min1440, names_to = "activity_minutes", values_to = "activity") 

hour_activity$activity_minutes = as.numeric(gsub("min", "", hour_activity$activity_minutes)) # remove "min" in the xlabel, and convert it to dbl

ggplot(hour_activity, aes(x = activity_minutes, y = activity, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(.~ education) +
  labs(
    title = "24-hour activity course",
    x = "time",
    y = "activity"
  ) +
  scale_x_continuous(limits = c(0,1440), 
                     breaks = seq(0,1440,150) 
                     ) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Male and female smooth lines are similar in group regardless of educational status, with no discernible differences. After 300 minutes, the action always becomes more intense. There is no discernible difference in the level of exercise difficulty between male and female participants with less than or equal to a high school education. The male participants, however, tend to be more active between 500 and 1200 minutes, whereas the female participants tend to be more active between 500 and 1200 minutes.
Participants often become more active during the day (400–1300 minutes) and less active after midnight (before 400 minutes). Participants with educational backgrounds below high school or with high school equivalents have an activity strength that is not gender-specific, but participants with educational backgrounds above high school have a different activity strength. Male participants with higher education levels tend to be more active in the morning and after dinner (for a total of about 500 minutes and 1200 minutes, respectively), whereas female participants with higher education levels likely to be more active just in the morning (for a total of about 500 minutes).
